Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal

### 2020-09-05 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
finished paper

To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the points of maximum absolute amplitude of their respective pulses. The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F_+ $; similarly, $ F_- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}. To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.

@@

% Key results in details

@@

% Comparison with results in other works
Direct comparison with other algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.

>>>>

To do so we will define the positive (upper) frontier and negative (lower) frontier as the set of points $ (i, |W[i]|) $ that are, necessarily but not sufficiently, the points of maximum absolute amplitude of their respective pulses. The set of all these points such that $ W[i] > 0 $ will be called the positive, piecewise linear frontier and will be denoted by $ F^+ $; similarly, $ F^- $ will denote the negative frontier of $ W $, defined by the subset of points $ (i, W[i]) $ where $ W[i] < 0 $, as is shown in figure \ref{fig:Frontiers}. To identify such points, we need to address the different units in which the amplitude and horizontal distance of our wave are expressed.

@@

Besides, the algorithm presents a definition of the position of the cycles of a wave, with which on can, for example, derive its fundamental frequency and phase: using one of the two frontiers, we have $ f = (\#F^\pm - 1) n / (F^\pm[-1] - F^\pm[0])$, where $ \#F^\pm $ denotes the cardinality of the frontier, while $ F^\pm[0] $ and $ F^\pm[-1] $ stand for the frontier's first and last item, respectively.

Similarly, each item of each of the frontiers can be used to estimate the local phase at that location. For the positive frontier, it is necessary to note that, for each item, the following equation holds: $ \cos(\phi + 2 \pi f i / n) = 1 $, and thus $ \phi = k 2 \pi - 2 \pi f i / n $, with $ k \in \mathbb{N } $, while for the negative frontier $ \cos(\phi + 2 \pi f i / n) = -1 $ and $ \phi = k 3 \pi - 2 \pi f i / n) $.

@@

\subsection{Comparison with traditional algorithms}

Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.

Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as cane be seen in figure \ref{fig:Comparison}.

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
  \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
  \label{fig:Comparison}
\end{figure}

For a recording of a guitar bend, we contrast the proposed algorithm with four common approaches of varied complexity: simple smoothing of the rectified original wave, lowpass filtering followed by peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artefacts shown in figure \ref{fig:AnalyticSignal}.

All those methods demanded careful parameter tuning: in the case of filtering, a local cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned. For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.

In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, that were also smoothed with the same procedure described. Figure \ref{fig:FrontiersToEnvelope} illustrates this process.

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
  \caption{Original positive and negative frontiers in dark gray and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
  \label{fig:FrontiersToEnvelope}
\end{figure}

Table \ref{table:Comparison} presents some computational results. The times are for custom implementations of those methods in the Python programming language. The least mean squares was calculated halving the envelopes presented in figure \ref{fig:Comparison} and comparing then with the rectified absolute wave, so as to avoid the potential influence of an estimated frequency. We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive. The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a mean to improve the efficiency of the algorithm.

\begin{table}[ht!]
\centering
\begin{tabular}{ l c c }
\hline
Method & LMS & time(s) \\
\hline
Presented Algorithm & 0.0088 & 4.1507 \\ 
Smoothing           & 0.0123 & 0.1911 \\ 
Lowpass Filter      & 0.0258 & 0.5601 \\ 
Hilbert Transform   & 0.0141 & 0.2599 \\ 
\hline
\end{tabular}
\caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
\label{table:Comparison}
\end{table}
